CrypTool Script \\ Mathematics and Cryptography

### 2002-09-27 script/de/chapters/numbertheory.tex
no message

Eine hervorragende "Ubersicht "uber die Rekorde im Faktorisieren zusammengesetzter Zahlen \index{Faktorisierungsrekorde}
mit unterschiedlichen Methoden findet sich auf der Webseite 
\href{http://www.crypto-world.com}{\texttt{http://www.crypto-world.com}}. Der aktuelle Rekord
(Stand Juni 2002) mit der GNFS-Methode (General Number Field Sieve) \index{General Number Field Sieve (GNFS)}
liegt in der Zerlegung einer 158-stelligen Zahl in ihre beiden Primfaktoren (diese haben 73 und 86 Dezimalstellen).

Dieser Rekord, aufgestellt am 18. Januar 2002 von Forschern der Universit"at Bonn \linebreak[4]
\href{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html}
{\texttt{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html}}, fand deutlich 
weniger Aufmerksamkeit in der Presse als die L"osung der RSA-Challenge vom 22. Oktober 1999, als niederl"andische
Forscher eine 155-stellige Zahl in ihre beiden 78-stellige Primfaktoren zerlegten (vergleiche Kapitel \ref{chptSecurityParam}).

@@

Die 6 kleineren, schon vorher gefundenen Primfaktoren dieser Zahl waren:
$$
\begin{array}{c}
        3, 1907, 425796183929, \\ 
        1624700279478894385598779655842584377, \\
        3802306738549441324432139091271828121 \quad{\rm und} \\ 
        128064886830166671444802576129115872060027.
\end{array}
$$
\begin{sloppypar}
Wobei die drei kleinsten Faktoren leicht (z.B. mit CrypTool unter Men"u {\bf Einzelverfahren / RSA-Demonstration / Faktorisieren}) bestimmt werden.
Die n"achsten drei Primfaktoren wurden von P.~Zimmerman
(\href{http://www.loria.fr/~zimmerma/ecmnet}{\tt http://www.loria.fr/~zimmerma/ecmnet}), 
T.~Grandlund
(\href{http://www.swox.se/gmp/}{\tt http://www.swox.se/gmp/})
und R. Harley in den Jahren 1999 und 2000 mit der Methode der elliptischen
Kurven gefunden.
\end{sloppypar}

@@

>>>>

Eine hervorragende "Ubersicht "uber die Rekorde im Faktorisieren zusammengesetzter Zahlen \index{Faktorisierungsrekorde}
mit unterschiedlichen Methoden findet sich auf der Webseite 
\href{http://www.crypto-world.com}{\texttt{http://www.crypto-world.com}}. Der aktuelle Rekord
(Stand Juni 2002) mit der GNFS-Methode (General Number Field Sieve) \index{General Number Field Sieve (GNFS)}
liegt in der Zerlegung einer $158$-stelligen Zahl in ihre beiden Primfaktoren (diese haben 73 und 86 Dezimalstellen).

Dieser Rekord, aufgestellt am 18. Januar 2002 von Forschern der Universit"at Bonn \linebreak[4]
\href{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html}
{\texttt{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html}}, fand deutlich 
weniger Aufmerksamkeit in der Presse als die L"osung der RSA-Challenge vom 22. Oktober 1999, als niederl"andische
Forscher eine $155$-stellige Zahl in ihre beiden $78$-stellige Primfaktoren zerlegten (vergleiche Kapitel \ref{chptSecurityParam}).
Das liegt unter anderem daran, dass die $512$ bit Zahl RSA-155 eine {\em magische} Grenze war.

@@

Die 6 kleineren, schon vorher gefundenen Primfaktoren dieser Zahl waren:
$$
\begin{array}{c}
        3, 1907, 425796183929, \\ 
        1624700279478894385598779655842584377, \\
        3802306738549441324432139091271828121 \quad{\rm und} \\ 
        128064886830166671444802576129115872060027.
\end{array}
$$
\begin{sloppypar}
Wobei die drei kleinsten Faktoren leicht (z.B. mit CrypTool unter Men"u {\bf Einzelverfahren / RSA-Demonstration / Faktorisieren}) bestimmt werden.
Die n"achsten drei Primfaktoren wurden von P.~Zimmerman
(\href{http://www.loria.fr/zimmerma/ecmnet}{\tt http://www.loria.fr/\~{}zimmerma/ecmnet}), 
T.~Grandlund
(\href{http://www.swox.se/gmp/}{\tt http://www.swox.se/gmp/})
und R. Harley in den Jahren 1999 und 2000 mit der Methode der elliptischen
Kurven gefunden.
\end{sloppypar}

@@

\begin{minipage}{\textwidth}
\vspace{3ex}
Verweise:
\vspace{-10pt}
\begin{itemize}
\item[]	  \href{http://www.loria.fr/\~{}zimmerma/records/gnfs158}{\texttt{http://www.loria.fr/\~{}zimmerma/records/gnfs158}} \\
          \href{http://www.crypto-world.com/FactorRecords.html}{\texttt{http://www.crypto-world.com/FactorRecords.html}} \\
	  \href{http://www.crypto-world.com/announcements/c158.txt}{\texttt{http://www.crypto-world.com/announcements/c158.txt}}
\end{itemize}
\end{minipage}

### 2002-09-27 script/en/chapters/numbertheory.tex
no message

$O(l)$ is the average number of processor steps depending on the bit length
$l$ of the number $n$ to be factorized.  For the best currently known
factorization algorithm the constant $c = (64/9)^/173) = 1,923$.

@@

This record, established on January 18, 2002 by researchers of the German
university of Bonn (see the press statement (in german language): \\
\href{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html}
{\texttt{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html}}), got much less attention within the press than the
solution of the RSA challenge, established on August 22, 1999, when
researchers from the Netherlands factorized the 155-digit number into its
both 78-digit primes (see Chapter \ref{chptSecurityParam}).

@@

The 6 smaller prime factors, already found before have been:
$$
\begin{array}{c}
3, 1907, 425796183929, \\
1624700279478894385598779655842584377, \\
3802306738549441324432139091271828121 \quad {\rm and} \\
128064886830166671444802576129115872060027.
\end{array}
$$
The first 3 factors can be easily computed (e.g. using CrypTool via menu {\bf Indiv. Procedures / RSA Demonstration / Factorise}).
The next three prime factors were found by P.~Zimmerman
(\href{http://www.loria.fr/~zimmerma/ecmnet}{\tt http://www.loria.fr/~zimmerma/ecmnet}), 
T.~Grandlund (\href{http://www.swox.se/gmp/}{\tt http://www.swox.se/gmp/})
and R. Harley during the years 1999 and 2000 using the elliptic curve factorisation method.

The last remaining factor, called ``C158'', was known to be composite by
then, but its factors where not known (the following 3 lines contain one number):
$$
\begin{array}{c}
39505874583265144526419767800614481996020776460304936 \\
45413937605157935562652945068360972784246821953509354 \\
4305870490251995655335710209799226484977949442955603
\end{array}
$$
The factorisation of C158 resulted in the following two prime factors:
$$
\begin{array}{c}
3388495837466721394368393204672181522815830368604993048084925840555281177
\end{array}
$$
and
$$
\begin{array}{c}
1165882340667125990314837655838327081813101 \\
2258146392600439520994131344334162924536139.
\end{array}
$$
So now all 8 prime factors of $2^{953} - 1$ have been found.
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

From a {\tt Mod} object you can get back the components with
the \texttt{component} or \texttt{lift}function:

>>>>

$O(l)$ is the average number of processor steps depending on the bit length
$l$ of the number $n$ to be factorized.  For the best currently known
factorization algorithm the constant $c = (64/9)^{1/173} = 1,923$.

@@

This record, established on January 18, 2002 by researchers of the German
university of Bonn (see the press statement (in german language): \\
\href{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html}
{\texttt{http://www.uni-bonn.de/Aktuelles/Pressemitteilungen/pm02/pm035-02.html}}), got much less attention within the press than the
solution of the RSA challenge, established on August 22, 1999, when
researchers from the Netherlands factorized the 155-digit number into its
both 78-digit primes (see Chapter \ref{chptSecurityParam}). 
One reason is that 512 bit RSA-155 was a kind of {\em magic} border.

@@

The 6 smaller prime factors, already found before have been:
$$
\begin{array}{c}
3, 1907, 425796183929, \\
1624700279478894385598779655842584377, \\
3802306738549441324432139091271828121 \quad {\rm and} \\
128064886830166671444802576129115872060027.
\end{array}
$$
The first 3 factors can be easily computed (e.g. using CrypTool via menu {\bf Indiv. Procedures / RSA Demonstration / Factorise}).
The next three prime factors were found by P.~Zimmerman
(\href{http://www.loria.fr/~zimmerma/ecmnet}{\tt http://www.loria.fr/\~{}zimmerma/ecmnet}), 
T.~Grandlund (\href{http://www.swox.se/gmp/}{\tt http://www.swox.se/gmp/})
and R. Harley during the years 1999 and 2000 using the elliptic curve factorisation method.

The last remaining factor, called ``C158'', was known to be composite by
then, but its factors where not known (the following 3 lines contain one number):
$$
\begin{array}{c}
39505874583265144526419767800614481996020776460304936 \\
45413937605157935562652945068360972784246821953509354 \\
4305870490251995655335710209799226484977949442955603
\end{array}
$$
The factorisation of C158 resulted in the following two prime factors:
$$
\begin{array}{c}
3388495837466721394368393204672181522815830368604993048084925840555281177
\end{array}
$$
and
$$
\begin{array}{c}
1165882340667125990314837655838327081813101 \\
2258146392600439520994131344334162924536139.
\end{array}
$$
So now all 8 prime factors of $2^{953} - 1$ have been found.

\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item[]	  \href{http://www.loria.fr/~zimmerma/records/gnfs158}{\texttt{http://www.loria.fr/~zimmerma/records/gnfs158}} \\
          \href{http://www.crypto-world.com/FactorRecords.html}{\texttt{http://www.crypto-world.com/FactorRecords.html}} \\
	  \href{http://www.crypto-world.com/announcements/c158.txt}{\texttt{http://www.crypto-world.com/announcements/c158.txt}}
\end{itemize}
\end{minipage}

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

From a {\tt Mod} object you can get back the components with
the \texttt{component} or \texttt{lift} function:
